Predicting the PCG PSeudo-Random Number Generator In Practice

### 2020-05-28 article/main.tex
retouche bib
tableaux flottants sections 3
Algos section 4 et 5

\usepackage[utf8]{inputenc}
\usepackage{amssymb, amsmath, amsfonts, amscd}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{url}
\usepackage{xspace}
\usepackage{subcaption}
%\usepackage{algorithm2e}
\usepackage{tikz}
\usepackage{cellspace}
\usepackage{multirow}
%\usepackage{parskip}
\usetikzlibrary{patterns}

@@

\usepackage[noend]{algpseudocode}
%\usepackage{algorithmicx}

@@

\begin{tabular}{|c|c|}
	\hline
	\(\ell\)  & minimum \(n\) \\
	\hline
	11 & 39 \\
	12 & 38 \\
	13 & 38 \\
	\hline
\end{tabular}

@@

\begin{tabular}{|c|c|c|c|}
	\hline
	n & \(\lVert G(n,64) \rVert \times \lVert G(n,64)^{-1} \rVert\) & \( \lambda_1(L(n,64)) \) & minimum \(\ell\) \\
	\hline
	3 & 2.869065991049632 & \(4.08909120094543e^{12}\) & 19 \\
	4 & 2.060717188416003 & \(2.43569932844650e^{14}\) & 13 \\
	5 & 3.7692142154392747 & \(1.71669732658884e^{15}\) & 11 \\
	6 & 2.6925925483728537 & \(1.02550589786192e^{16}\) & 8 \\
	\hline
\end{tabular}

@@

Therefore, the algorithm that reconstructs the internal state of the
\textsf{PCG64} generator with known increment proceeds as follows:

  \begin{algorithmic}[1]
    \Procedure{ReconstructState${}_\ell$}{$X_0, X_1, X_2$}
    \For{$0 \leq w < 2^{\ell}$} \Comment{Guess least-significant bits of $S_0$}
    \State $w_j = a^j w \bmod 2^{\ell}$ \Comment{Propagate to $S_1, S_2$}
    \State $S'_j \gets a^j w + c \sum_{k = 0}^j a^k \bmod 2^{128}$    \Comment{Shifted states}

@@

    \State $ S_1 \gets a  S_0 + c$ \Comment{Check consistency of $X_1$}
    \State $\widehat{Y}_1 =  S_1[0:64] \oplus  S_1[64:128]$

    \If{$\widehat {Y}_1 = Y_1$}
    \State output $ S_0$ as a candidate internal state.
    \EndIf
    \EndFor
    \EndFor
    \EndProcedure
  \end{algorithmic}

\begin{theorem}
  Algorithm \textsc{ReconstructState} is correct (it always return the correct internal state).
\end{theorem}

\todo{avec la nouvelle partie 3 et le blabla au dessus de l'algo, y'a plus vraiment besoin de preuve}

\section{Predicting PCG with Unknown Increment}
\label{sec:Cunknown}

The same algorithm does not apply directly to the general case where the value
of $c$ is unknown. The global strategy to obtain a prediction algorithm is
essentially the same (target a truncated version of the underlying linear
congruential generator), but is a bit more complicated to implement.

Set $\Delta S_i = S_{i+1} - S_i$; it is easily checked that $\Delta S_i$ is a
geometric progression of common ratio $a$. Therefore, reconstructing both $S_0$
and $\Delta S_0$ is sufficient to compute all subsequent states (and recover the
unknown increment $c$). We note that
\begin{align*}
  S_i - S_0 \equiv \sum_{j=1}^i S_j - S_{j-1} \equiv \sum_{j=0}^{i-1} \Delta S_j \equiv \Delta S_0 \cdot \sum_{j=0}^{i-1} a^j \equiv \Delta S_0 \frac{a^i-1}{a-1} \mod 2^{128}
\end{align*}

The prediction algorithm proceeds in three phases:
\begin{enumerate}
\item Reconstruct $\Delta S_0[0:64+\ell]$ from $X_0, \dots, X_{4}$, check consistency with $X_5, \dots, X_{39}$.
\item Fully reconstruct $\Delta S_0$ from this partial knowledge and the outputs.
\item Reconstruct $S_0$ from $\Delta S_0$ and the outputs.
\end{enumerate}

@@

\subsection{Phase 1: Partial Difference Reconstruction}

In order to access to a part of $\Delta S_i$, we use the same strategy as in
section~\ref{sec:Cknown}: we guess the least significant bits of $S_0$ and some
rotations. The difference is that, in addition, to obtain the ``long-term
advantage'' in the least significant bits of the states, we must also guess the
least significant bits of $c$ as well ($c$ is always odd; this makes one less
bit to guess). We must also guess one additional rotation to compute the
differences $\Delta S_i$.

Another problem is that of confirming that the guesses are correct. When $c$ is
known, we could reconstruct the internal state; from there, filtering out the
bad guesses is easy. When $c$ is unknown, a very strong consistency check can be
implemented as follows: once we have found $\Delta S_0[0:64+\ell]$, we can
compute $\Delta S_i[0:64+\ell]$ for any $i$; because we have guessed the first
rotation and $\ell$ lest significant bits of the state, using~\eqref{eq:output}
we gain access to $S_0[58:64+\ell]$; combined with the ``differences'', this
reveals $S_i[58:64+\ell]$ for any $i$ (and we already had $S_i[0:\ell]$). This
means that we can compute $S_i[0:\ell] \oplus S_i[64:64+\ell]$ for any $i$.

Given a ``fresh'' output $X_i$, and assuming that the guesses are correct, then we should have:
\begin{equation}\label{eq:consistency}
  S_i[0:\ell] \oplus S_i[64:64+\ell] \in \bigl\{ (X_i \lll r_i)[0:\ell]~|~r_i=0, 1, \dots, 63 \bigr\}.
\end{equation}
If none of the rotations yields a match, then the guesses made beforehand have
to be wrong. As soon as $\ell \geq 6$, this test is efficient (and it can be
done for any number of fresh $X_i$'s). As a consequence, bad guesses can be
filtered deterministically.

\begin{algorithmic}[1]
  \Procedure{ConsistencyCheck}{$\Delta S_0, w_0, Y_0, X_5, \dots, X_k$}
  \State // \emph{Implements the test given by eq.~\eqref{eq:consistency}}.
  \State $x_0 = w_0 \oplus Y_0[0:\ell]$ \Comment{$x_0 = S_0[64:64+\ell]$}
  \For{$i=5, \dots, k$}
  \State $x_i = x_0 + \left(\Delta S_0 \frac{a^i-1}{a-1} \bmod 2^{64+\ell}\right)[64:64+\ell]$. \Comment{$x_i = S_i[64:64+\ell]$}
%  \State $\Delta S_i[\ell:64+\ell] \gets a^i \Delta U_0 \bmod 2^{64}$
  \State $\mathrm{Target} \gets S_i[0:\ell] \oplus x_i$
  \If{$\mathrm{Target} \not\in \{ (X_i \lll r_i)[0:\ell]~|~ 0\leq r_i < 64\}$}
  % \State $\mathrm{Found} \gets \mathbf{False}$
  % \For{$0 \leq r_i < 64$}
  % \State $Y_i \gets X_i \lll r_i$
  % \If{$Y_i[0:\ell] = \mathrm{Target}$}
  % \State $\mathrm{Found} \gets \mathbf{True}$ 
  % \EndIf
  % \EndFor
  % \If{not Found}
  \State \Return \textbf{False} \Comment{Bad Guesses}
  \EndIf
  \EndFor
  \State \Return \textbf{True} \Comment{Good  Guesses}
  \EndProcedure

@@

\Procedure{ReconstructPartialDifference}{$X_0, \dots, X_k$}
  \For{$0 \leq w_0 < 2^{\ell}$ and $0 \leq c < 2^{\ell - 1}$} \Comment{Guess least-significant bits}
  \State $S'_j \gets a^j w_0 + (2c+1) \sum_{k = 0}^j a^k \bmod 2^{128}$ \Comment{Shifted states}
  \For{$0 \leq r_0, r_1, r_2, r_3, r_4 < 64$} \Comment{Guess rotations}
  \State $Y_j \gets X_j \lll r_j$ \Comment{Undo rotations}
  \State $T_j \gets \left(r_j \oplus Y_j[58:64]\right) +  64 \cdot \left(w_j \oplus Y_j[0:\ell]\right)$ \Comment{Truncated LCG}
  \State $T'_j \gets (T_j -  S'_i[58:64+\ell]) \bmod 2^{6+\ell}$ \Comment{Shifted truncated LCG}
  \State $\Delta T'_j = (T'_{j+1} - T'_j) \bmod 2^{6 + \ell}$  \Comment{Difference}
  \State $(\Delta U_0, \dots, \Delta U_3) \gets \left\lfloor (\Delta T'_0, \dots, \Delta T'_3) \cdot 2^{58-\ell} \cdot \widetilde G_4^{-1} \right\rceil \cdot \widetilde G_4$ \Comment{CVP}
  \State $\Delta S_0[0:64+\ell] \gets \left((w_1 - w_0) \bmod 2^{\ell}\right) + 2^{\ell} \cdot \Delta U_0[0:58-\ell]$ \Comment{Check}
  \If{$\Call{ConsistencyCheck}{\Delta_0, w_0, Y_0, X_5, \dots, X_k}$}
  \State \Return $(w_0, c, r_0, \dots, r_4, \Delta_0)$.
  \EndIf
  \EndFor
  \EndFor
  \EndProcedure
\end{algorithmic}

\begin{theorem}
  If $\ell \geq 13$, then algorithm \textsc{ReconstructPartialDifference} is
  correct (it always returns the correct values).
\end{theorem}

\begin{proof}
  \todo{WRITE-ME}
  \begin{align*}
  \mathbf{S} &= \LCG(S_0, c) \\
             &= \LCG\left(S_0[0:\ell] + 2^\ell S_0[\ell:128], c[0:\ell] + 2^\ell c[\ell:128]\right) \\
             &= \LCG\left(S_0[0:\ell], c[0:\ell]\right) + 2^\ell \LCG\left(S_0[\ell:128],  c[\ell:128]\right) \\
  \Delta \mathbf{S} &= \LCG(S_1, c) - \LCG(S_0, c) \\
             &= \LCG(S_1 - S_0, 0)
\end{align*}
\end{proof}

@@

\section{Implementation of the Predictor and Practical Results}

>>>>

\usepackage[utf8]{inputenc}
\usepackage{amssymb, amsmath, amsfonts, amscd}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{url}
\usepackage{xspace}
\usepackage{subcaption}
\usepackage{algorithm}
\usepackage{tikz}
\usepackage{cellspace}
\usepackage{multirow}
%\usepackage{parskip}
\usetikzlibrary{patterns}

@@

\usepackage[noend]{algpseudocode}
%\usepackage{algorithmic}

@@

\begin{table}
  \centering
    \begin{tabular}{|c|c|}
	\hline
	\(\ell\)  & minimum \(n\) \\
	\hline
	11 & 39 \\
	12 & 38 \\
	13 & 38 \\
	\hline
\end{tabular}
  \caption{\todo{WRITE-ME}}
  \label{tab:exact_cvp}
\end{table}

@@

\begin{table}
  \centering
  \begin{tabular}{|c|c|c|c|}
	\hline
	n & \(\lVert G(n,64) \rVert \times \lVert G(n,64)^{-1} \rVert\) & \( \lambda_1(L(n,64)) \) & minimum \(\ell\) \\
	\hline
	3 & 2.869065991049632 & \(4.08909120094543e^{12}\) & 19 \\
	4 & 2.060717188416003 & \(2.43569932844650e^{14}\) & 13 \\
	5 & 3.7692142154392747 & \(1.71669732658884e^{15}\) & 11 \\
	6 & 2.6925925483728537 & \(1.02550589786192e^{16}\) & 8 \\
	\hline
  \end{tabular}
  \caption{\todo{WRITE-ME}}
  \label{tab:babai}
\end{table}

@@

Therefore, the algorithm that reconstructs the internal state of the
\textsf{PCG64} generator with known increment proceeds as shown in algorithm~\ref{algo:known}.

\begin{algorithm}
\begin{algorithmic}[1]
    \Procedure{ReconstructState${}_\ell$}{$X_0, X_1, X_2$}
    \For{$0 \leq w < 2^{\ell}$} \Comment{Guess least-significant bits of $S_0$}
    \State $w_j = a^j w \bmod 2^{\ell}$ \Comment{Propagate to $S_1, S_2$}
    \State $S'_j \gets a^j w + c \sum_{k = 0}^j a^k \bmod 2^{128}$    \Comment{Shifted states}

@@

    \State $ S_1 \gets a  S_0 + c$ \Comment{Recompute $X_1$}
    \State $\widehat{Y}_1 =  S_1[0:64] \oplus  S_1[64:128]$

    \If{$\widehat {Y}_1 = Y_1$} \Comment{Check consistency}
    \State output $ S_0$ as a candidate internal state.
    \EndIf
    \EndFor
    \EndFor
    \EndProcedure
  \end{algorithmic}
  \caption{State reconstruction Algorithm (case where $c$ is known)}
  \label{algo:known}
\end{algorithm}

% \begin{theorem}
%   Algorithm \textsc{ReconstructState} is correct (it always return the correct internal state).
% \end{theorem}

% \todo{avec la nouvelle partie 3 et le blabla au dessus de l'algo, y'a plus vraiment besoin de preuve}
% Charles: je suis plutôt d'accord.

\section{Predicting PCG64 with Unknown Increment}
\label{sec:Cunknown}

The algorithm of section~\ref{sec:Cknown} does not apply directly to the general
case where the value of $c$ is unknown. A ``baseline'' procedure would consist
in guessing $S_0[64:128]$ and $S_1[64:128]$; using eq.~\eqref{eq:output}, this
would reveal $S_0$ and $S_1$, from which the increment $c$ could be easily
deduced. This would take $2^{128}$ iterations of a very simple procedure, but is
completely infeasible.

The global strategy to obtain a practical prediction algorithm is essentially
the same as before (target a truncated version of the underlying linear
congruential generator), but is a bit more complicated to implement.

Set $\Delta S_i = S_{i+1} - S_i$; it is easily checked that $\Delta S_i$ is a
geometric progression of common ratio $a$. Therefore, reconstructing both $S_0$
and $\Delta S_0$ is sufficient to compute all subsequent states (and recover the
unknown increment $c$). Let us set:
\begin{align*}
  \nabla S_i \stackrel{def}{=} S_i - S_0 \equiv \sum_{j=1}^i S_j - S_{j-1} \equiv \sum_{j=0}^{i-1} \Delta S_j \equiv \Delta S_0 \cdot \sum_{j=0}^{i-1} a^j \equiv \Delta S_0 \frac{a^i-1}{a-1} \mod 2^{128}
\end{align*}
Note that $\nabla S_0 = \Delta S_0$. The prediction algorithm we propose proceeds in three phases:
\begin{enumerate}
\item Reconstruct $\Delta S_0[0:64+\ell]$ from $X_0, \dots, X_{4}$, check consistency with $X_5, \dots, X_{39}$.
\item Fully reconstruct $\Delta S_0$ from this partial knowledge and the outputs.
\item Reconstruct $S_0$ from $\Delta S_0$ and the outputs.
\end{enumerate}

@@

\subsection{Partial Difference Reconstruction}

In order to access to a part of $\Delta S_i$, we use the same
``guess-and-determine'' strategy as in section~\ref{sec:Cknown}: we guess the
least significant bits of $S_0$ and some rotations, then check consistency. The
difference is that, since $c$ is unknown, we must in addition guess the least
significant bits of $c$ to obtain the same ``long-term advantage'' ($c$ is
always odd; this makes one less bit to guess). We must also guess $k+1$
successive rotation to get information on $k$ successive differences
$\Delta S_i$.

Another problem is that of confirming that the guesses are correct. When $c$ was
known, we could reconstruct the internal state; from there, filtering out the
bad guesses was easy. When $c$ is unknown, the same strategy does not work, but
a very strong consistency check can still be implemented. The essential idea is
the following: once $\nabla S_0[0:64+\ell]$ is known, then all the rotations can
be determined accurately and with little computational effort. This in turn
allows a consistency check between the partial information deduced from the
guesses and the actual outputs.

Once we have found $\nabla S_0[0:64+\ell]$, we can compute
$\nabla S_i[0:64+\ell]$ for any $i$ (geometric progression); because we have
guessed the first rotation and the $\ell$ least significant bits of the state,
using~\eqref{eq:output} we gain access to $S_0[58:64+\ell]$; combined with the
``differences'' $\nabla S_i$, this reveals $S_i[58:64+\ell]$ for any $i$ (and we
already had $S_i[0:\ell]$). This allows us to compute
$S_i[0:\ell] \oplus S_i[64:64+\ell]$ for any $i$.

Given a ``fresh'' output $X_i$, and assuming that the guesses are correct, then we should have:
\begin{equation}\label{eq:find_rotation}
  S_i[0:\ell] \oplus S_i[64:64+\ell] = (X_i \lll r_i)[0:\ell].
\end{equation}
If $X_i$ is known, then this gives a way to find the $i$-th rotation $r_i$ by
trying the 64 possible values. As soon as $\ell \geq 6$, this should determine
$r_i$ almost uniquely. In any case, if the guesses were correct, then we should have:
\begin{equation}\label{eq:consistency}
  S_i[0:\ell] \oplus S_i[64:64+\ell] \in \bigl\{ (X_i \lll r)[0:\ell]~|~0 \leq r < 64 \bigr\}.
\end{equation}
If none of the 64 possible rotations yields a match, then the guesses made
beforehand have to be wrong. As a consequence, bad guesses can be filtered
deterministically.

More precisely, let us assume that we have guessed the $\ell$ least-significant
bits of $S_0$ (we denote them by $w_0$) and the first rotation $r_0$. Set
$Y_0 \stackrel{def}{=} X_0 \lll r_0$; there is an unknown value
$x_0$ such that:
\[
  S_0 = w_0 + 2^\ell x + 2^{64} (w_0 \oplus Y_0[0:\ell]) + 2^{64+\ell} (x \oplus Y_0[\ell:64]) \qquad \left(0 \leq x < 2^{64-\ell}\right).
\]
The unknown value $x_0$ will not be recovered until the third phase of the whole
procedure. We obtain the $i$-th state by
$S_i \equiv \nabla S_i + S_0 \bmod 2^{64+\ell}$; however, because the ``middle''
of $S_0$ is unknown, an unknown \emph{carry} may cross the 64-th during the
addition and perturb $S_i[64:64+\ell]$. As a result, there are unknown
quantities $0 \leq y_i < 2^{64-\ell}$ and $0 \leq z_i \leq 1$ such that:
\begin{align*}
  S_i[0:\ell] &\equiv \nabla S_i + w_0 \bmod 2^\ell \\
  S_i[\ell:64] &\equiv y_i \\
  S_i[64:64+\ell] &\equiv z_i + \nabla S_i[64:64+\ell] + (w_0 \oplus Y_0[0:\ell]) \bmod 2^\ell \\
\end{align*}

In algorithm~\ref{algo:unknown_1}, \textsc{ConsistencyCheck} uses
eq.~\eqref{eq:consistency} combined with this observation to discard bad guesses.

\begin{algorithm}
\begin{algorithmic}[1]
  \Procedure{ConsistencyCheck}{$\Delta S_0, w_0, Y_0, X_5, \dots, X_k$}
  \State $v_0 = w_0 \oplus Y_0[0:\ell]$ \Comment{$v_0 = S_0[64:64+\ell]$}
  \For{$i=5, \dots, k$}
  \State $u_i \gets \Delta S_0 (a^i-1)(a-1)^{-1} \bmod 2^{64+\ell}$ \Comment{$u_i = \nabla S_i[0:64+\ell]$}
  \State $w_i = w_0 + u_i[0:\ell] \bmod 2^{\ell}$ \Comment{$w_i = S_i[0:\ell]$}
  \State $v_i = v_0 + u_i[64:64+\ell] \bmod 2^{\ell}$ \Comment{$S_i[64:64+\ell] \in \{v_i, v'_i\}$}
  \State $v'_i = v_i + 1 \bmod 2^{\ell}$
  \State $\mathcal{C} \gets \{ w_i \oplus (X_i \lll r_i)[0:\ell]~|~ 0\leq r_i < 64\}$ \Comment{Check eq.~\eqref{eq:consistency}}
  \If{$\{v_i, v'_i\} \cap \mathcal{C} = \emptyset$}
  \State \Return \textbf{False} \Comment{Bad Guesses}
  \EndIf
  \EndFor
  \State \Return \textbf{True} \Comment{No inconsistency}
  \EndProcedure

@@

\Procedure{ReconstructPartialDifference}{$X_0, \dots, X_k$}
  \For{$0 \leq w_0 < 2^{\ell}$ and $0 \leq c < 2^{\ell - 1}$} \Comment{Guess least-significant bits}
  \State $S'_j \gets a^j w_0 + (2c+1) \sum_{k = 0}^j a^k \bmod 2^{128}$ \Comment{Shifted states}
  \For{$0 \leq r_0, r_1, r_2, r_3, r_4 < 64$} \Comment{Guess rotations}
  \State $Y_j \gets X_j \lll r_j$ \Comment{Undo rotations}
  \State $T_j \gets \left(r_j \oplus Y_j[58:64]\right) +  64 \cdot \left(w_j \oplus Y_j[0:\ell]\right)$ \Comment{Truncated LCG}
  \State $T'_j \gets (T_j -  S'_i[58:64+\ell]) \bmod 2^{6+\ell}$ \Comment{Shifted truncated LCG}
  \State $\Delta T'_j = (T'_{j+1} - T'_j) \bmod 2^{6 + \ell}$  \Comment{Difference}
  \State $(\Delta U_0, \dots, \Delta U_3) \gets \left\lfloor (\Delta T'_0, \dots, \Delta T'_3) \cdot 2^{58-\ell} \cdot \widetilde G_4^{-1} \right\rceil \cdot \widetilde G_4$ \Comment{CVP}
  \State $\Delta S_0[0:64+\ell] \gets \left((w_1 - w_0) \bmod 2^{\ell}\right) + 2^{\ell} \cdot \Delta U_0[0:58-\ell]$ \Comment{Check}
  \If{$\Call{ConsistencyCheck}{\Delta_0, w_0, Y_0, X_5, \dots, X_k}$}
  \State \Return $(w_0, c, r_0, \dots, r_4, \Delta_0)$.
  \EndIf
  \EndFor
  \EndFor
  \EndProcedure
\end{algorithmic}
\caption{Partial difference reconstruction algorithm (when $c$ is unknown).}
\label{algo:unknown_1}
\end{algorithm}

The heart of the algorithm is again the reconstruction of a truncated geometric
progression. Looking at table~\ref{tab:babai}, we see that the best choice
consists in guessing 5 consecutive rotations and $\ell=13$ least-significant
bits. Therefore, \textsc{ReconstructPartialDifference} does $2^{55}$ iterations
of the inner loop, and succeeds deterministically.

% \begin{theorem}
%   If $\ell \geq 13$, then algorithm \textsc{ReconstructPartialDifference} is
%   correct (it always returns the correct values).
% \end{theorem}

% \begin{proof}
%   \todo{WRITE-ME}
%   \begin{align*}
%   \mathbf{S} &= \LCG(S_0, c) \\
%              &= \LCG\left(S_0[0:\ell] + 2^\ell S_0[\ell:128], c[0:\ell] + 2^\ell c[\ell:128]\right) \\
%              &= \LCG\left(S_0[0:\ell], c[0:\ell]\right) + 2^\ell \LCG\left(S_0[\ell:128],  c[\ell:128]\right) \\
%   \Delta \mathbf{S} &= \LCG(S_1, c) - \LCG(S_0, c) \\
%              &= \LCG(S_1 - S_0, 0)
% \end{align*}
% \end{proof}

\subsection{Full Difference Reconstruction}

Knowing the values of $\Delta S_0[0:64+\ell]$ as well as the $\ell$
least-significant bits of $S_0$ and $c$ is sufficient to get rid the nastier
feature \textsf{PCG64}: armed with this knowledge, we can determine all the
subsequent rotations deterministically, at negligible cost, using
eq~\eqref{eq:find_rotation}.

For each $i$, it suffices to try the $64$ possible values of $r_i$; only one
should satisfy the equation. Using $X_0, \dots, X_{40}$, we thus recover the 6
most-significant bits of $S_0, \dots, S_{40}$. We therefore compute the 6 most
significant bits of the differences $\Delta S_i$ between consecutive states (up
to missing carries), and we are faced with the problem of reconstructing a
128-bit geometric progression using 39 consecutive outputs truncated to their 6
most-significant bits. This can be done by solving an instance of CVP in
dimension $\leq 40$. We use the off-the-shelf CVP solver embedded in
\texttt{fplll}: it runs in no measurable time.

More precisely, there are still details to be dealt with regarding
carries. Indeed, the \textsc{ReconstructRotations} returns the 6 most
significant bits of any number of $S_i$'s (denoted by $r_i$). Therefore, there
are unknown 122-bit quantities $x_i$ such that:
\begin{align*}
  S_i &= 2^{122} r_i + x_i \\
  \Delta S_i &\equiv 2^{122} (r_{i+1} - r_i) + (x_{i+1} - x_{i}) \bmod 2^{128}
\end{align*}

\begin{algorithm}
\begin{algorithmic}[1]
  \Function{ReconstructRotations}{$\Delta S_0, v_0, i, k$}
 \State // \emph{Return a list of potential} $\texttt{[}r_i, r_{i+1}, \dots, r_k\texttt{]}$; \emph{assume that $v_0 = S_0[64:64+\ell]$}
  \If{$i > k$}
  \State \Return \texttt{[]} \Comment{End recursion}
  \EndIf
  \State $\mathcal{T} \gets \Call{ReconstructRotations}{\Delta S_0, v_0, i + 1, k}$ \Comment{Find all the $(r_{i+1}, \dots, r_{k})$}
  \State $\mathcal{H} \gets \texttt{[]}$ \Comment{List of possible $r_i$'s}

  \State $u_i \gets \Delta S_0 (a^i-1)(a-1)^{-1} \bmod 2^{64+\ell}$ \Comment{$u_i = \nabla S_i[0:64+\ell]$}
  \State $w_i = w_0 + u_i[0:\ell] \bmod 2^{\ell}$ \Comment{$w_i = S_i[0:\ell]$}
  \State $v_i = v_0 + u_i[64:64+\ell] \bmod 2^{\ell}$ \Comment{$S_i[64:64+\ell] \in \{v_i, v'_i\}$}
  \State $v'_i = v_i + 1 \bmod 2^{\ell}$

  %\State $x_i = x_0 + \left(\Delta S_0 \frac{a^i-1}{a-1} \bmod 2^{64+\ell}\right)[64:64+\ell]$. \Comment{$x_i = S_i[64:64+\ell]$}
  \For{$0 \leq r < 64$} \Comment{Try all rotations}
  \If{$w_i \oplus (X_i \lll r)[0:\ell] \in \left\{v_i, v'_i\right\}$} \Comment{Check eq.~\eqref{eq:find_rotation}}
  \State $\mathcal{H} \gets r \texttt{::} \mathcal{H}$ \Comment{New candidate $r_i$}
  \EndIf
  \EndFor
  \State \Return $\left\{ h\texttt{::} t~|~ h \in \mathcal{H}, t \in \mathcal{T} \right\}$ \Comment{Return $\mathcal{H} \times \mathcal{T}$}
  \EndFunction
  \State
  \Function{ReconstructFullDifference}{...}
  \State \todo{WRITEME}
  \EndFunction
\end{algorithmic}
\caption{Rotations and full difference reconstruction algorithm}
\end{algorithm}

@@

\section{Implementation and Practical Results}
